Forecasting the old-age dependency ratio to determine the best pension age

### 2019-01-13 PensionAge.tex
Draft_13/Jan/2019

\usepackage{todonotes}

@@

where \(m_{x,t}\) is the central death
rate\footnote{The central death rate is properly defined in Section~\ref{sec:data}}
for age \(x\) in year \(t\). This model involves fitting a matrix of log
central rates with sets of age-specific constants, \{\(a_x\)\} and
\{\(b_x\)\}, and a time-varying index \(k_t\). \(b_x\) is a set of
parameters that tell us the rate at which each age responds to changes
in \(k_t\). \(k_t\) measures the general level of the log death rates;
\(a_x\) is the general pattern across age of the log mortality rate and
\(\epsilon_{x,t}\) is an error term with zero mean and constant
variance, reflecting the randomness that can't be captured by the model.

The model is under-determined in terms of estimation of parameters. For
example, if the vector \(a,b,k\) is one solution, then for any scalar
\(c, a-bc, b, k+bc\) also must be one solution. Therefore, Lee and
Carter (1992) impose the following constraints:

\begin{equation}
  \sum_{t=1}^{n}k_t=0, \sum_{x=x_1}^{x_m}b_x=1
\end{equation}

@@

\subsubsection{Hyndman-Ullah Method}\label{hyndman-ullah-method}

@@

The model proposed by \textcite{HU07} is as follows:

@@

Equation \eqref{eq:HU_1} describes the smoothing process. \(y_t(x)\) is
the observed data in year \(t\) and age \(x\), and \(s_t(x)\) is the
smoothed function of \(y_t(x)\), estimated by constrained weighted
penalized regression splines \autocite{BHT14}. \(\sigma_t(x)\) allows
the variance to be a function of time and age; and \(\epsilon_{t,x}\)
are independent identical distributed standard normal error terms.
Equation \eqref{eq:HU_2} has a similar structure to the LC method,
except that it incorporates more than one principal component and score.
It describes the dynamics of \(s_t(x)\) over time. The \(\beta_{t,j}\)
for every \(j\) and \(\mu(x)\) are sets of age-specific constants, and
\(\phi_j(x)\) for every \(j\) is a time-varying index. The \(e_t(x)\) is
the model error, which is the residual function with mean zero and no
serial correlation. The \(\mu(x)\) is estimated by the average of
\(s_t(x)\) across years, which measures the general shape of the
observed data \(y_t(x)\). The estimates of \(\phi_j(x)\) and
\(\beta_{t,j}\) are the \(j\)th principal component and scores of
\(s_t(x)-\mu(x)\) respectively. In other words, we take singular value
decomposition of \(s_t(x)-\mu(x)\), then \eqref{eq:HU_2} now becomes

\begin{equation*}
  s_t(x)-\mu(x) = \phi_1(x)\beta_{t,1}+\phi_2(x)\beta_{t,2}+\dots+\phi_k(x)\beta_{t,k}
\end{equation*}

@@

In terms of forecasting, by conditioning on the observed data
\(\bm{I}=\left\{{y_1(x),\dots,y_n(x)}\right\}\) and the set of
functional principal components
\(\bm{B}=\left\{{\phi_1(x),\phi_2(x),\dots,\phi_j(x)}\right\}\), the
\(h\)-step-ahead forecast of log rate \(y_{n+h}(x)\) can be acquired by:

@@

where \(\widehat{\beta}_{n+h\mid n,j}\) denotes the \(h\)-step-ahead
forecast of \(\beta_{n+h\mid n,j}\). Because of the time-varying
property of \(\left\{\beta_{t,j}\right\}\), it controls the dynamics of
the process. For every component series \(\left\{\beta_{t,j}\right\}\),
we fit them with an ARIMA\((p_j,d_j,q_j)\) process, where
\(p_j,d_j,q_j\) are determined by minimizing the Akaike Information
Criterion (Booth, Hyndman \& Tickle, 2014). The fitted process is given
as follows:

@@

We use the same notation and similar procedures to construct data as in
\textcite{HB08}, because the fundamental model for forecasting
population for this study is the \textcite{HU07} method. The data used
to construct population age structure included: age- and gender-specific
birth and death numbers of each calendar year, age- and gender-specific
population members on 1~January of each year, and age- and
gender-specific exposures to risk (i.e., population of age \(x\) at 30
June) for each year. The notations are as follows:

@@

where \(x\) denotes age and \(x=0,1,2,\dots,m^+\). \(m^+\) is the
open-ended upper age group, which is 100 in this study. \(t\) denotes
year and \(t=1,2,\dots,n\), where \(n\) is the horizon of historical
data used. Therefore, all of those data points can be recorded in an
\(m\times n\) matrix. The super-scripts \(M\) and \(F\) denote male and
female respectively.

We can obtain historical mortality rates \(m_t(x)\) (also called central
death rates) and fertility rates \(f_t(x)\) as follows:

@@

In addition, the net
migration\footnote{Net migration is the difference between immigration and emigration in a certain area during a specified time frame}
(denoted \(G\)) is estimated using a demographic growth-balance
equation, which is an equation to summarize the relationship between
population change and three factors: net migration, births and deaths.
The equation is expressed as: \todo{Please define $p$ and $^+$}

@@

where \(G_t(x,x+1)\) refers to net migration in calendar year \(t\) of
persons aged \(x\) at beginning of year \(t\), \(G_t(p-1^+,p)\) refers
to migration in calendar year \(t\) of persons aged \(p-1\) and older at
beginning of year \(t\), and \(G_t(B,0)\) refers to deaths in calendar
year \(t\) of births during year \(t\); and similarly for deaths,
\(D_t(x,x+1), D_t(p-1^+,p^+), D_t(B,0)\). The deaths are estimated using
the standard life table approach of population projection. It should be
noted that the estimated net migration includes errors in data
recording.

Historical central death rates and start-year and mid-year populations
of residents in Australia by sex and age in single years for age group
0-99 and 100+ have been sourced from the \textcite{HMD} for 1921--2014.
The start-year population for 2015 is also used to estimate net
migration. Data for age-specific annual fertility rates by single years
of age for 15-49 over the period 1950-2014 was obtained from the
\textcite{ABS12}. Even though annual mortality rates are available from
1921 to 2009, only data collected after 1950 is used for two reasons.
First, the period between 1921-1950 has very different mortality
patterns from later years because of wars and epidemics \autocite{HB08}.
Since the model for this study is valid under the assumption that
extrapolative methods can be used, it seemed best to delete the less
relevant data, which might influence estimations of parameters if we
include them. Secondly, fertility rates are only available from 1950, so
the mortality rate was cut to fit the same period.
\todo{Han has updated the data period.}

@@

In this study, three steps were taken to simulate OADR associated with a
specific pension age scheme. A stochastic functional model was applied
to five components of population change: female mortality rates, male
mortality rates, female net migration, male net migration and fertility
rates to simulate these five components over the prediction horizon
(which is 50 years in this study). age- and gender-specific population
paths over the prediction horizon were simulated using the demographic
growth-balance equation. The Monte Carlo simulation method was used
consistently in this research. Finally, calculation of OADR associated
with a specific pension age scheme was performed by applying the OADR
formula. Even though the process seems complicated, it is relatively
resy to implement using \textit{R}. The detailed explanation of each
step is as follows. The corresponding R code is attached in Appendix B.
\todo{There is no appendix, Han recommends to put the R code as a supplement}

@@

The main procedure for forecasting the population age structure follows
\textcite{HB08}, except that the data is updated, and we incorporate the
product-ratio method \autocite{HBY13} to improve the accuracy of the
forecast. The \textcite{HU07} model, introduced in section 2.3.2, is
applied to the mortality rate, fertility rate and net migration
separately, based on the assumption that they behave independently.
Therefore, to fit the model to historical fertility rates, let
\(y_t(x)\) in equation \eqref{eq:HU_1} denote the observed historical
fertility data. However, for mortality rate by sex, the product-ratio
method is used to ensure the forecast of females and males are
non-divergent or coherent \autocite{HBY13}. In the product-ratio method,
after smoothing the female and male mortality rates, two other series
were created: product series (denoted by \(p_t(x)\)) and ratio series
(denoted by \(r_t(x)\)), which are defined as follows:

\begin{align}
  p_t(x)&=\sqrt{s_{t,M}(x)s_{t,F}(x)}, \label{eq:coherent_1}\\
  r_t(x)&=\sqrt{\frac{s_{t,M}(x)}{s_{t,F}(x)}}. \label{eq:coherent_2}
\end{align}

Then, \textcite{HU07} was applied to these two new variables instead of
male and female mortality rate as follows:

@@

where time series \(\left\{\beta_{t,j}\right\}\) are fitted with an
ARIMA model while \(\left\{\gamma_{t,l}\right\}\) are fitted by
stationary ARFIMA model (i.e., \(d\) in ARIMA(\(p,d,q\)) constrained to
\(0<d<\frac{1}{2}\)). It is the stationary constraint on
\(\gamma_{t,l}\) that ensures the forecast of male and female do not
diverge. Then, simulation h-step-ahead forecast values of product
variable and ratio variable was performed (denoted by
\(p_{n+h\mid n}(x)\) and \(r_{n+h\mid n}(x)\) respectively) according to
\eqref{eq:coherent_1} and \eqref{eq:coherent_2}. From here, obtaining
h-step-ahead forecast of sex-specific coherent mortality rates (denoted
\(m_{n+h\mid n,M}(x)\) and \(m_{n+h\mid n,F}(x)\)) was performed using a
back-transformation:

\begin{align*}
  m_{n+h\mid n,M}(x) & = p_{n+h\mid n}(x)r_{n+h\mid n}(x)\\
  m_{n+h\mid n,F}(x) & = p_{n+h\mid n}(x)/r_{n+h\mid n}(x)
\end{align*}

@@

Even though the OADR is a simple function of the population age
structure and pension age, the formula can't be directly applied to the
simulated population array because the pension age changes over time. To
overcome this problem, we assume that the pension age can only be
adjusted at the beginning of a year for these calculations. Therefore,
one pension age is applicable for a whole year and hence OADR is
generated at the beginning of each year to show the trend. We will show
the process of obtaining a one-step-ahead OADR (i.e., forecast of OADR
in year 2011), given a specific pension age. The same method can be
applied to all years over the prediction horizon, given their
corresponding pension age. Further, the fractional pension age causes
practical issues in calculating OADR, since ages are recorded as
integers. We will also show how this issue was handled in the example
below.

@@

\begin{equation*}
  \bm{p}_{n+1}=[p_{m,i}] =
    \begin{bmatrix}
       p_{1,1} & \cdots\cdots & p_{1,1000} \\
      \vdots & \ddots & \vdots \\
      p_{100^+,1} & \cdots\cdots &p_{100^+,1000}
      \end{bmatrix}.
\end{equation*}

@@

When \(r\neq0\), the vector of total number of working age population in
\eqref{eq:work} is added, and the vector of total number of aged
population in \eqref{eq:aged} is subtracted with an adjustment:

\begin{equation}
  \text{adjustment}=0.r\times100%[p_{a,1},p_{a,2},\dots,p_{a,1000}]
\end{equation}

This adjustment is based on the assumption that the birthday of
population is uniformly distributed over the year. Then,
\(r\times100\%\) of population with a recorded integer age of \(a\) at
beginning of the year is actually younger than \(a.r\), so they are
still under the pension age. In other words, \((1-r\times100\%)\) of
population with recorded integer age \(a\) at beginning of the year is
actually older than \(a.r\), so they should be counted as the aged
population. The simulation vectors of total work age population and
total aged population should then be:
\todo{Han feel the second equation does not have \%?}

\begin{align*}
  \bm{p}^{\text{work}}_{n+1}
    & = \left[p^{\text{work}}_{n+1,1},p^{\text{work}}_{n+1,2},\dots,p^{\text{work}}_{n+1,1000}\right]\\
    & = \left[\sum_{m=15}^{a-1}p_{m,1},\sum_{m=15}^{a-1}p_{m,2},\dots,\sum_{m=15}^{a-1}p_{m,1000}\right]+0.r\times100\%\left[p_{a,1},p_{a,2},\dots,p_{a,1000}\right]   \\
  \bm{p}^{\text{aged}}_{n+1}
    & = \left[p^{\text{aged}}_{n+1,1},p^{\text{aged}}_{n+1,2},\dots,p^{\text{aged}}_{n+1,1000}\right]\\
    & = \left[\sum_{m=a}^{100^+}p_{m,1},\sum_{m=a}^{100^+}p_{m,2},\dots,\sum_{m=a}^{100^+}p_{m,1000}\right]-0.r\times100%\left[p_{a,1},p_{a,2},\dots,p_{a,1000}\right]
\end{align*}

After the adjustment, we can generate simulation vectors of the
one-step-ahead OADR by dividing every element of
\(\bm{p}^{\text{work}}_{n+1}\) by the corresponding element of
\(\bm{p}^{\text{aged}}_{n+1}\), i.e.:

@@

\begin{equation}
  \bm{\text{OADR}}^{forecast}=\left[\text{OADR}_{h,i}\right] =
  \begin{bmatrix}
    \text{OADR}_{n+1,1} & \cdots\cdots & \text{OADR}_{n+50,1000} \\
    \vdots & \ddots & \vdots \\
    \text{OADR}_{n+50,1} & \cdots\cdots & \text{OADR}_{n+50,1000}
  \end{bmatrix}.
\end{equation}

@@

We use the fitted model of fertility rates as an example to interpret
the results of fitted models. It should be noted that models fitted to
product and ratio series are less likely to be interpreted in an
economic sense. Figure \ref{fig:componentsfertility} above shows the
mean age pattern (\(\mu(x)\)), the first two basis functions
(\(\phi_1(x)\) and \(\phi_2(x)\)) and first two coefficients (i.e.,
principal component scores \(\beta_{t,1}\) and \(\beta_{t,2}\)) for
fertility rates. The first term (\(\phi_1(x)*\beta_{t,1}\)) accounts for
the most variation in fertility, so a relatively sensible interpretation
can be given to the first term. In the plot of coefficient 1, estimated
historical first principal component scores \(\beta_{t,1}\) are shown in
black for 1950-2009. It captures the general trend of fertility change
over time. It indicates a sharp increase between 1950-1960, followed by
a rapid decline of fertility rates from 1961. This increase of fertility
is linked to the Post-World War II baby
boom\footnote{Between the years 1946 and 1964} in Australia, resulting
from the recovery from war and the Great
Depression\footnote{The Great Depression was a severe economic depression in the decade before World War II}.
The dramatic decrease from 1960 was due to the availability of
contraception pills in Australia. A noticeable further reduction of
fertility rates in late 1960s may be due to a change in the law allowing
abortion in some states in Australia.

\begin{figure}
\centering
\includegraphics{PensionAge_files/figure-latex/componentsfertility-1.pdf}
\caption{\label{fig:componentsfertility}Fitted basis functions and
coefficients for Australian fertility rates.}
\end{figure}

The first basis function \(\phi_1(x)\) measures the different effects of
the trend over time across ages. It indicates the general pattern of
fertility has very little effect on females of age 15. This makes
intuitive sense because most early pregnancies are accidental,
therefore, any of previously discussed events has no effect on early
pregnancies. In addition, there are two spikes in the first basic
function, which suggests that the historical trend has a distinct impact
on females aged around their 20s and 40s. Furthermore, the blue line is
the fitted value of future coefficients (i.e., \(\beta_{n+h,1}\), for
\(h=1,2,\dots,50\)) and the yellow shading shows their 80\% prediction
interval according to the fitted ARIMA model of first coefficients
\(\left\{\beta_{t,1}\right\}\). Other terms further explain variation
that has not been captured by first term, but less economic explanations
can be given for them.

\begin{figure}
\centering
\includegraphics{PensionAge_files/figure-latex/totalpopulationbyyears-1.pdf}
\caption{\label{fig:totalpopulationbyyears}Fifth-year forecasts of total
population for each sex, along with 80\% prediction intervals. The
actually population from 1950 to 2015 is shown using circles.}
\end{figure}

\begin{figure}
\centering
\includegraphics{PensionAge_files/figure-latex/totalpopulationin2028-1.pdf}
\caption{\label{fig:totalpopulationin2028}Forecast population pyramid for
2028, along with 80\% prediction intervals. The actual pyramid for 2010
is shown using dashed lines.}
\end{figure}

Both Figures \ref{fig:totalpopulationbyyears} and
\ref{fig:totalpopulationin2028} show the forecast results of population.
Figure \ref{fig:totalpopulationbyyears} shows the forecast of total
population by gender over the prediction horizon, while Figure
\ref{fig:totalpopulationin2028} shows the forecast of population age
structure in a specific year, 2028. \ref{fig:totalpopulationbyyears}
indicates that both the female and male population are expected to
increase to 18 million by 2060. Since the fertility rate is expected to
remain low, the possible reason of the significant increase in
population size is net migration. We can also observe that the further
the projected year is, the bigger the forecast variation of population.
This increase in variation shows that we have less relevant information
for further apart prediction horizons.

Figure \ref{fig:totalpopulationin2028} shows the mean and the 80\%
prediction interval of simulated population paths for year 2028, along
with the 2010 base population. It can be observed that overall
population size is expected to increase substantially. The uncertainty
of population of infant is much higher than other ages, which reflects
the higher forecast variation of fertility rates than mortality rates.
The uncertainty of the young age population is mainly due to migration
\autocite{HB08}.

\begin{figure}
\centering
\includegraphics{PensionAge_files/figure-latex/OADRat65-1.pdf}
\caption{\label{fig:OADRat65}Fifty-year forecasts of the OADR with
assumption of pension age equal to 65 for both sexes, along with 80\%
prediction interval.}
\end{figure}

\begin{figure}
\centering
\includegraphics{PensionAge_files/figure-latex/OADRcurrent-1.pdf}
\caption{\label{fig:OADRcurrent}Fifty-year forecasts of the OADR associated
with a pension age equal to 65 (black line) and equal to the pension age
scheme with approved change (red line) respectively, along with the
desired OADR level in grey (i.e., mean forecast of OADR in 2014). The
actual OADRs for 1950-2010 are shown in black.}
\end{figure}

>>>>

\usepackage{todonotes, xcolor}

@@

where \(m_{x,t}\) is the central death
rate\footnote{The central death rate is properly defined in Section~\ref{sec:data}}
for age \(x\) in year \(t\). This model involves fitting a matrix of log
central death rates with sets of age-specific constants, \{\(a_x\)\} and
\{\(b_x\)\}, and a time-varying index \(k_t\). \(b_x\) is a set of
parameters that tell us the rate at which each age responds to changes
in \(k_t\). \(k_t\) measures the general level of the log death rates;
\(a_x\) is the general pattern across age of the log mortality rate and
\(\epsilon_{x,t}\) is an error term with zero mean and constant
variance, reflecting the randomness that can't be captured by the model.

The model is under-determined in terms of estimation of parameters. For
example, if the vector \(a,b,k\) is one solution, then for any scalar
\(c\), \(a, b/c, kc\) or \(a - bc, b, k+c\) also must be two possible
solutions. Therefore, Lee and Carter (1992) impose the following
constraints:

\begin{equation}
  \sum_{t=1}^{n}k_t=0, \qquad \sum_{x=x_1}^{x_m}b_x=1
\end{equation}

@@

\subsubsection{\texorpdfstring{Hyndman-Ullah Method
\label{sec:HU}}{Hyndman-Ullah Method }}\label{hyndman-ullah-method}

@@

The model proposed by \textcite{HU07} is as follows: \vspace{-.3in}

@@

Equation \eqref{eq:HU_1} describes the smoothing process. \(y_t(x)\) is
the observed data in year \(t\) and age \(x\), and \(s_t(x)\) is the
smoothed function of \(y_t(x)\), estimated by constrained weighted
penalized regression splines \autocite{BHT14}. \(\sigma_t(x)\) allows
the variance to be a function of time and age; and \(\epsilon_{t,x}\)
are independent identical distributed standard normal error terms.
Equation \eqref{eq:HU_2} has a similar structure to the LC method,
except that it incorporates more than one principal component and score.
It describes the dynamics of \(s_t(x)\) over time. The \(\beta_{t,j}\)
for every \(j\) and \(\mu(x)\) are sets of age-specific constants, and
\(\phi_j(x)\) for every \(j\) is a time-varying index. The \(e_t(x)\) is
the model error, which is the residual function with mean zero and no
serial correlation. The \(\mu(x)\) is estimated by the average of
\(s_t(x)\) across years, which measures the general shape of the
observed data \(y_t(x)\). The estimates of \(\phi_j(x)\) and
\(\beta_{t,j}\) are the \(j\)th principal component and scores of
\(s_t(x)-\mu(x)\) respectively. In other words, we take singular value
decomposition of \(s_t(x)-\mu(x)\), then \eqref{eq:HU_2} now becomes
\vspace{-.2in}

\begin{equation*}
  s_t(x)-\mu(x) = \phi_1(x)\beta_{t,1}+\phi_2(x)\beta_{t,2}+\dots+\phi_J(x)\beta_{t,J}
\end{equation*}

@@

In terms of forecasting, by conditioning on the observed data
\(\bm{I}=\left\{{y_1(x),\dots,y_n(x)}\right\}\) and the set of
functional principal components
\(\bm{B}=\left\{{\phi_1(x),\phi_2(x),\dots,\phi_j(x)}\right\}\), the
\(h\)-step-ahead forecast of log rate \(y_{n+h}(x)\) can be acquired by:
\vspace{-.2in}

@@

where \(\widehat{\beta}_{n+h\mid n,j}\) denotes the \(h\)-step-ahead
forecast of \(\beta_{n+h\mid n,j}\). Because of the time-varying
property of \(\left\{\beta_{t,j}\right\}\), it controls the dynamics of
the process. For every component series \(\left\{\beta_{t,j}\right\}\),
we fit them with an ARIMA\((p_j,d_j,q_j)\) process, where
\(p_j,d_j,q_j\) are determined by minimizing the Akaike Information
Criterion (Booth, Hyndman \& Tickle, 2014). The fitted process is given
as follows: \vspace{-.3in}

@@

We use the same notation and similar procedures to construct data as in
\textcite{HB08}, because the fundamental model for forecasting
population for this study is the \textcite{HU07} method. The data used
to construct population age structure included: age- and gender-specific
birth and death numbers of each calendar year, age- and gender-specific
population members on 1~January of each year, and age- and
gender-specific exposures to risk (i.e., population of age \(x\) at 30
June) for each year. The notations are as follows: \vspace{-.4in}

@@

\vspace{-.4in}where \(x\) denotes age and \(x=0,1,2,\dots,m^+\). \(m^+\)
is the open-ended upper age group, which is 100 in this study. \(t\)
denotes year and \(t=1,2,\dots,n\), where \(n\) is the horizon of
historical data used. Therefore, all of those data points can be
recorded in an \(m\times n\) matrix. The superscripts M and F denote
male and female respectively.

We can obtain historical mortality rates \(m_t(x)\) (also called central
death rates) and fertility rates \(f_t(x)\) as follows: \vspace{-.3in}

@@

In addition, the net
migration\footnote{Net migration is the difference between immigration and emigration in a certain area during a specified time frame}
(denoted \(G\)) is estimated using a demographic growth-balance
equation, which is an equation to summarize the relationship between
population change and three factors: net migration, births and deaths.
The equation is expressed as: \todo{Please define $p$ and $^+$}
\vspace{-.3in}

@@

where \(G_t(x,x+1)\) refers to net migration in calendar year \(t\) of
persons aged \(x\) at beginning of year \(t\), \(G_t(p-1^+,p)\)
\todo{should it be $p^+$} refers to migration in calendar year \(t\) of
persons aged \(p-1\) and older at beginning of year \(t\), and
\(G_t(B,0)\) refers to deaths in calendar year \(t\) of births during
year \(t\); and similarly for deaths,
\(D_t(x,x+1), D_t(p-1^+,p^+), D_t(B,0)\). The deaths are estimated using
the standard life table approach of population projection. It should be
noted that the estimated net migration includes errors in data
recording.

Historical central death rates and start-year and mid-year populations
of residents in Australia by sex and age in single years for age group
0-99 and 100+ have been sourced from the \textcite{HMD} for
1921--2014.\todo{Han has updated the data period.} The start-year
population for 2015 is also used to estimate net migration. Data for
age-specific annual fertility rates by single years of age for 15-49
over the period 1950-2014 was obtained from the \textcite{ABS12}. Even
though annual mortality rates are available from 1921 to 2014, only data
collected after 1950 is used for two reasons. First, the period between
1921-1950 has very different mortality patterns from later years because
of wars and epidemics \autocite{HB08}. Since the model for this study is
valid under the assumption that extrapolative methods can be used, it
seemed best to delete the less relevant data, which might influence
estimations of parameters if we include them. Secondly, fertility rates
are only available from 1950, so the mortality rate was cut to fit the
same period.

@@

In this study, three steps were taken to simulate OADR associated with a
specific pension age scheme. A stochastic functional data model was
applied to five components of population change: female mortality rates,
male mortality rates, female net migration, male net migration and
fertility rates to simulate these five components over the prediction
horizon (which is 50 years in this study). Age- and gender-specific
population paths over the prediction horizon were simulated using the
demographic growth-balance equation. The Monte Carlo simulation method
was used consistently in this research. Finally, calculation of OADR
associated with a specific pension age scheme was performed by applying
the OADR formula. Even though the process seems complicated, it is
relatively resy to implement using \textit{R} \autocite{Team18}. The
detailed explanation of each step is as follows. The corresponding R
code is attached as a supplement.

@@

The main procedure for forecasting the population age structure follows
\textcite{HB08}, except that the data is updated, and we incorporate the
product-ratio method \autocite{HBY13} to improve the accuracy of the
forecast. The \textcite{HU07} model, introduced in section \ref{sec:HU},
is applied to the mortality rate, fertility rate and net migration
separately, based on the assumption that they behave independently.
Therefore, to fit the model to historical fertility rates, let
\(y_t(x)\) in equation \eqref{eq:HU_1} denote the observed historical
fertility data. However, for mortality rate by sex, the product-ratio
method is used to ensure the forecast of females and males are
non-divergent or coherent \autocite{HBY13}. In the product-ratio method,
after smoothing the female and male mortality rates, two other series
were created: product series (denoted by \(p_t(x)\)) and ratio series
(denoted by \(r_t(x)\)), which are defined as follows: \vspace{-.4in}

\begin{align}
  p_t(x)&=\sqrt{s_{t}^{\text{M}}(x)s_{t}^{\text{F}}(x)}, \label{eq:coherent_1}\\
  r_t(x)&=\sqrt{\frac{s_{t}^{\text{M}}(x)}{s_{t}^{\text{F}}(x)}}. \label{eq:coherent_2}
\end{align}

Then, \textcite{HU07} was applied to these two new variables instead of
male and female mortality rate as follows: \vspace{-.3in}

@@

where time series \(\left\{\beta_{t,j}\right\}\) are fitted with an
ARIMA model while \(\left\{\gamma_{t,l}\right\}\) are fitted by
stationary ARFIMA model (i.e., \(d\) in ARIMA(\(p,d,q\)) constrained to
\(0<d<\frac{1}{2}\)). It is the stationary constraint on
\(\gamma_{t,l}\) that ensures the forecast of male and female do not
diverge. Then, simulating \(h\)-step-ahead forecast values of product
variable and ratio variable was performed (denoted by
\(p_{n+h\mid n}(x)\) and \(r_{n+h\mid n}(x)\) respectively) according to
\eqref{eq:coherent_1} and \eqref{eq:coherent_2}. From here, obtaining
\(h\)-step-ahead forecast of sex-specific coherent mortality rates
(denoted \(m_{n+h\mid n,M}(x)\) and \(m_{n+h\mid n,F}(x)\)) was
performed using a back-transformation: \vspace{-.3in}

\begin{align*}
  m_{n+h\mid n}^{\text{M}}(x) & = p_{n+h\mid n}(x)r_{n+h\mid n}(x)\\
  m_{n+h\mid n}^{\text{F}}(x) & = p_{n+h\mid n}(x)/r_{n+h\mid n}(x)
\end{align*}

@@

Even though the OADR is a simple function of the population age
structure and pension age, the formula can't be directly applied to the
simulated population array because the pension age changes over time. To
overcome this problem, we assume that the pension age can only be
adjusted at the beginning of a year for these calculations. Therefore,
one pension age is applicable for a whole year and hence OADR is
generated at the beginning of each year to show the trend. We will show
the process of obtaining a one-step-ahead OADR (i.e., forecast of OADR
in year 2015), given a specific pension age. The same method can be
applied to all years over the prediction horizon, given their
corresponding pension age. Further, the fractional pension age causes
practical issues in calculating OADR, since ages are recorded as
integers. We will also show how this issue was handled in the example
below.

@@

\begin{equation*}
  \bm{p}_{n+1}=[p_{n+1,m,i}] =
    \begin{bmatrix}
       p_{1,1} & \cdots\cdots & p_{1,1000} \\
      \vdots & \ddots & \vdots \\
      p_{100^+,1} & \cdots\cdots &p_{100^+,1000}
      \end{bmatrix}.
\end{equation*}

@@

When \(r\neq0\), the vector of total number of working age population in
\eqref{eq:work} is added, and the vector of total number of aged
population in \eqref{eq:aged} is subtracted with an adjustment:
\vspace{-.2in}

\begin{equation}
  \text{adjustment}=0.r\times100\%[p_{a,1},p_{a,2},\dots,p_{a,1000}]
\end{equation}

This adjustment is based on the assumption that the birthday of
population is uniformly distributed over the year. Then,
\(r\times100\%\) of population with a recorded integer age of \(a\) at
beginning of the year is actually younger than \(a.r\), so they are
still under the pension age. In other words, \((1-r\times100\%)\) of
population with recorded integer age \(a\) at beginning of the year is
actually older than \(a.r\), so they should be counted as the aged
population. The simulation vectors of total work age population and
total aged population should then be:

\begin{align*}
  \bm{p}^{\text{work}}_{n+1}
    & = \left[p^{\text{work}}_{n+1,1},p^{\text{work}}_{n+1,2},\dots,p^{\text{work}}_{n+1,1000}\right]\\
    & = \left[\sum_{m=15}^{a-1}p_{m,1},\sum_{m=15}^{a-1}p_{m,2},\dots,\sum_{m=15}^{a-1}p_{m,1000}\right]+0.r\times100\%\left[p_{a,1},p_{a,2},\dots,p_{a,1000}\right]   \\
  \bm{p}^{\text{aged}}_{n+1}
    & = \left[p^{\text{aged}}_{n+1,1},p^{\text{aged}}_{n+1,2},\dots,p^{\text{aged}}_{n+1,1000}\right]\\
    & = \left[\sum_{m=a}^{100^+}p_{m,1},\sum_{m=a}^{100^+}p_{m,2},\dots,\sum_{m=a}^{100^+}p_{m,1000}\right]-0.r\times100\%\left[p_{a,1},p_{a,2},\dots,p_{a,1000}\right]
\end{align*}

After the adjustment, we can generate simulation vectors of the
one-step-ahead OADR by dividing every element of
\(\bm{p}^{\text{work}}_{n+1}\) by the corresponding element of
\(\bm{p}^{\text{aged}}_{n+1}\), i.e.: \vspace{-.3in}

@@

\begin{equation*}
  \bm{\text{OADR}}^{\text{forecast}}=\left[\text{OADR}_{h,i}\right] =
  \begin{bmatrix}
    \text{OADR}_{n+1,1} & \cdots\cdots & \text{OADR}_{n+1,1000} \\
    \vdots & \ddots & \vdots \\
    \text{OADR}_{n+50,1} & \cdots\cdots & \text{OADR}_{n+50,1000}
  \end{bmatrix}.
\end{equation*}

@@

We use the fitted model of fertility rates as an example to interpret
the results of fitted models. It should be noted that models fitted to
product and ratio series are less likely to be interpreted in an
economic sense. Figure \ref{fig:componentsfertility} above shows the
estimated mean age pattern \(\widehat{\mu}(x)\), the first two estimated
basis functions, \(\widehat{\phi}_1(x)\) and \(\widehat{\phi}_2(x)\),
and first two estimated coefficients (i.e., estimated principal
component scores \(\widehat{\beta}_{t,1}\) and
\(\widehat{\beta}_{t,2}\)) for fertility rates. The first term
(\(\widehat{\phi}_1(x)\times \widehat{\beta}_{t,1}\)) accounts for the
most variation in fertility, so a relatively sensible interpretation can
be given to the first term. In the plot of the first coefficient,
estimated historical first principal component scores
\(\widehat{\beta}_{t,1}\) are shown in black for 1950-2014. It captures
the general trend of fertility change over time. It indicates a sharp
increase between 1950-1960, followed by a rapid decline of fertility
rates from 1961. This increase of fertility is linked to the Post-World
War II baby boom\footnote{Between the years 1946 and 1964} in Australia,
resulting from the recovery from war and the Great
Depression\footnote{The Great Depression was a severe economic depression in the decade before World War II}.
The dramatic decrease from 1960 was due to the availability of
contraception pills in Australia. A noticeable further reduction of
fertility rates in late 1960s may be due to a change in the law allowing
abortion in some states in Australia.

\begin{figure}
\centering
\includegraphics{PensionAge_files/figure-latex/componentsfertility-1.pdf}
\caption{\label{fig:componentsfertility}Fitted basis functions and
coefficients for Australian age-specific fertility rates.}
\end{figure}

The first estimated basis function \(\widehat{\phi}_1(x)\) measures the
different effects of the trend over time across ages. It indicates the
general pattern of fertility has very little effect on females of age
15. This makes intuitive sense because most early pregnancies are
accidental, therefore, any of previously discussed events has no effect
on early pregnancies. In addition, there are two spikes in the first
basic function, which suggests that the historical trend has a distinct
impact on females aged around their 20s and 40s. Furthermore, the blue
line is the fitted value of future coefficients (i.e.,
\(\widehat{\beta}_{n+h,1}\), for \(h=1,2,\dots,50\)) and the yellow
shading shows their 80\% pointwise prediction interval according to the
fitted ARIMA model of first coefficients
\(\left\{\widehat{\beta}_{t,1}\right\}\). Other terms further explain
variation that has not been captured by first term, but less economic
explanations can be given for them.

Both Figures \ref{fig:totalpopulationbyyears} and
\ref{fig:totalpopulationin2034} show the forecast results of population.
Figure \ref{fig:totalpopulationbyyears} shows the forecast of total
population by gender over the prediction horizon, while Figure
\ref{fig:totalpopulationin2028} shows the forecast of population age
structure in a specific year, 2034. Figure
\ref{fig:totalpopulationbyyears} indicates that both the female and male
population are expected to increase to around 23 million by 2064. Since
the fertility rate is expected to remain low, the possible reason of the
significant increase in population size is net migration. We can also
observe that the further the projected year is, the bigger the forecast
variation of population. This increase in variation shows that we have
less relevant information for further apart prediction horizons.

\begin{figure}
\centering
\includegraphics{PensionAge_files/figure-latex/totalpopulationbyyears-1.pdf}
\caption{\label{fig:totalpopulationbyyears}Fifty-year forecasts of total
population for each sex, along with 80\% prediction intervals. The
actually population from 1950 to 2015 is shown using circles.}
\end{figure}

\begin{figure}
\centering
\includegraphics{PensionAge_files/figure-latex/totalpopulationin2034-1.pdf}
\caption{\label{fig:totalpopulationin2034}Forecast population pyramid for
2034, along with 80\% pointwise prediction intervals. The actual pyramid
for 2015 is shown using dashed lines.}
\end{figure}

Figure \ref{fig:totalpopulationin2034} shows the mean and the 80\%
pointwise prediction interval of simulated population paths for year
2034, along with the 2015 base population. It can be observed that
overall population size is expected to increase substantially. The
uncertainty of population of infant is much higher than other ages,
which reflects the higher forecast variation of fertility rates than
mortality rates. The uncertainty of the young age population is mainly
due to migration \autocite{HB08}.

\begin{figure}
\centering
\includegraphics{PensionAge_files/figure-latex/OADRat65-1.pdf}
\caption{\label{fig:OADRat65}Fifty-year forecasts of the OADR with
assumption of pension age equal to 65 for both sexes, along with 80\%
pointwise prediction interval.}
\end{figure}

\begin{figure}
\centering
\includegraphics{PensionAge_files/figure-latex/OADRcurrent-1.pdf}
\caption{\label{fig:OADRcurrent}Fifty-year forecasts of the OADR associated
with a pension age equal to 65 (green line) and equal to the pension age
scheme with approved change (orange line) respectively, along with the
desired OADR level in purple. The actual OADRs for 1950-2014 are shown
in black.}
\end{figure}
